A Streaming Algorithm for 2-Center with Outliers in High Dimensions

### 2015-05-03 paper.tex
Another submission
Referring to appendix A inside the paper

%\paragraph{Streaming Model}
To implement our algorithm in the streaming model,
we maintain the sets $B_1$ and $B_2$ 
in a data structure that supports adding points,
and gives a $\beta$-approximation to 1-center with $k$ outliers, for $k=0,\dots,z$.
Moreover, we maintain a set $B_u = B_1 \cup B_2$ in a similar data structure.
Note that these data structures do not need to maintain all the points. 
They only need to have a buffer of size $\dz$ that keeps the most recently added points.

@@

\begin{theorem}
\label{thm:far-cmplx}
	If $\delta^* > \alpha r^*$, a $1.8$-approximation to the 2-center problem with $z$ outliers 
	can be maintained in $O(d^3z^2)$ space and $\poly(d,z)$ update time.
\end{theorem}

\begin{proof}
Our data structures for maintaining $B_1$, $B_2$, and $B_u$
employ the streaming algorithm of~\cite{zarrabi2009streaming,chan2014streaming},
%To maintain an approximation to 1-center with $z$ outliers,
%we use the streaming algorithm of~\cite{zarrabi2009streaming}
%which combined with the %current best streaming algorithm for 1-center~\cite{chan2014streaming},
%result of~\cite{chan2014streaming},
which has an approximation factor of $1.22 \times \sqrt{2} <1.8$.
The algorithm of~\cite{zarrabi2009streaming}
uses $O(d^3z)$ space and $\poly(d,z)$ update time,
which implies the same bounds for \Cref{alg:case-1}.
Since we need to run $z + 1$ instances of \Cref{alg:case-1} in parallel, 
the space and update time is multiplied by a factor of $z$.
\end{proof}

@@

\REM{ % to be added in the camera-ready version
\paragraph{Acknowledgement}
The authors would like to thank Kiana Ehsani and Sahand Mozaffari 
for their thoughtful discussions, 
and for their help to improve the quality of this paper.
}

@@

\section{Answering Queries}

%Using \Cref{alg:case-1},
In the following, we show how to answer queries of this kind:
find two $\alpha$-separated congruent balls of minimum radius
to cover all but at most $z$ points of the stream processed so far.

Our query algorithm is presented in \Cref{alg:query}.
The idea behind the algorithm is as follows.
By our initial assumption about $p_1$ and by invariants (c), 
we know that $B_1$ completely contains $B_1^*$.
%because $B_1 \cap B_2^* = \emptyset$ . 
But for $B_2$, it might be the case that our assumption about $c_2$ 
was incorrect, and therefore, $B_2$ may not completely contain $B_2^*$. 
The overcome this, we try all possible candidates for $c_2$,
and compute, for each possible combinations of 
$B_1$, $B_2$, and Buffer, the best solution for 2-center with $z$ outliers
using function \textproc{ComputeMin} presented in \Cref{alg:compute-min}.

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
	%\STATE solutions $\gets \set{\Call{ComputeMin}{B_1, B_2, \Buffer}}$
	\STATE solutions $\gets \set{}$

	\STATE candidates $\gets \set{c_2} \cup \Buffer $
	\IF {$\card{B_2} < \dz$}
		\STATE add $B_2$ to candidates
	\ELSE
		\STATE $B_1 \gets B_u$
	\ENDIF

	\STATE $\delta_0 = \delta$
	\FOR{$c \in$ candidates}
		\STATE $\rc  \gets \frac{2}{\alpha} \len{c_1 c}$
		\STATE $B'_1 \gets B_1$, $\delta \gets \max \{ \delta_0, r \}$
		\STATE $B'_2 \gets \emptyset$, $\Buffer' \gets \emptyset$ 
		\FOR{$p \in$ candidates}
			\IF {\NOT (\Call{AddToB$'_1$}{$p$} \OR \Call{AddToB$'_2$}{$p$})}
				\STATE add $p$ to $\Buffer'$
			\ENDIF
		\ENDFOR
		\STATE add \Call{ComputeMin}{$B'_1$, $B'_2$, $\Buffer'$} to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\sc ComputeMin($B'_1, B'_2, \text{\rm Buffer}$)} 
\algsetup{indent=1.5em}
\label{alg:compute-min}
\begin{algorithmic}[1]
%\Function{ComputeMin}{$B_1$, $B_2$, Buffer}
	\STATE solutions $\gets \set{}$
	\FOR{$k \gets 0, \dots,(z - \card{\mbox{Buffer}})$}
		\STATE $r_1 \gets \Call{1-Center}{B'_1, k}$
		\STATE $r_2 \gets \Call{1-Center}{B'_2, z - \card{\mbox{Buffer}} - k}$
		\STATE add $\max \set{r_1, r_2}$ to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

@@

\noindent
%Now, we show that the query algorithm correctly considers all possible candidates for $c_2$.
Let $C$ denote the set of candidates for $c_2$.
By invariant~(c), we know that $B_1 \cap B_2^* = \emptyset$. 
Therefore, there exists a point in $(B_2 \cup \Buffer) \cap B_2^*$,
and hence, $C$ is $(B_2 \cup \Buffer) \cap B_2^*$ in general.
%We now use the following lemma.
%In this general case, our candidates $C$ for $c_2$ is set $(B_2 \cup \Buffer) \cap B_2^*$.
However, when $|B_2| \geq \dz$, 
we know by Lemma \ref{lem:B2dz} that $(\{c_2\} \cup \Buffer) \cap B_2^* \neq \emptyset$. 
Therefore, if $|B_2| \ge \dz$, then our candidates for $C$ are only $\{c_2\} \cup \Buffer$.

Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals to current $c_2$, then, $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b1) and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we just need to see which points in $\Buffer \cap B_2$ are located in $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that assumed $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without contradicting invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} use functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} for adding a point to $B'_1$ and $B'_2$ respectively. 
The definition of these functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
except that, they add point to $B'_i$ instead of $B_i$ ($i = 1, 2$).

Since \Cref{alg:query} considers all valid candidates for $c$, 
at least for one $c^* \in C$, we have $c^* \in B_2^*$. 
We call the corresponding $B'_1$ and $B'_2$ by $B''_1$ and $B''_2$. 
Since by Observation \ref{obs:c+4},
$1 \leq \frac{\len{c_1 c^*}}{\alpha^*} < \frac{\len {c_1 c^*}}{\alpha r^*}$, 
we have $2r^* \leq \radius{c^*}$, and therefore,
$B^*_1 \subset B''_1$ and $B^*_2 \subset B''_2$ by Lemma~\ref{lem:2r}.
On the other hand, since the distance of the new points added to $B''_1$ compare to $B_1$ is less than $\len{c_1 p} \leq \max \{\delta, \radius{c^*} \}$, 
we have by invariant (c) that  $B''_1 \cap B_2^* = \emptyset$.
As a result $B''_1$ (resp., $B''_2$) completely covers $B_1^*$  (resp., $B_2^*$), 
and points in $\Buffer$ are certainly outliers.
The only unknown part is that \Cref{alg:query} does not know how many outliers are in $B''_1$ and $B''_2$.
Therefore, \Cref{alg:compute-min} tries all possible cases and choose the one with the minimum radius.

@@

>>>>

%\paragraph{Streaming Model}
To implement our algorithm in the streaming model,
we maintain the sets $B_1$ and $B_2$ 
in a data structure that supports adding points,
and gives a $\beta$-approximation to 1-center with $z$ outliers. %, for $k=0,\dots,z$.
Moreover, we maintain a set $B_u = B_1 \cup B_2$ in a similar data structure.
Note that these data structures do not need to maintain all the points. 
They only need to have a buffer of size $\dz$ that keeps the most recently added points.

@@

\begin{theorem}
\label{thm:far-cmplx}
	If $\delta^* > \alpha r^*$, a $1.8$-approximation to the 2-center problem with $z$ outliers 
	can be maintained in $O(d^3z^2)$ space and $\poly(d,z)$ update/query time.
\end{theorem}

\begin{proof}
Our algorithm for answering queries is provided in ~\Cref{sec:query}.
To maintain $B_1$, $B_2$, and $B_u$, we 
use the streaming algorithm of~\cite{zarrabi2009streaming,chan2014streaming},
%To maintain an approximation to 1-center with $z$ outliers,
%we use the streaming algorithm of~\cite{zarrabi2009streaming}
%which combined with the %current best streaming algorithm for 1-center~\cite{chan2014streaming},
%result of~\cite{chan2014streaming},
which has an approximation factor of $1.22 \times \sqrt{2} <1.8$.
%The algorithm of~\cite{zarrabi2009streaming}
The algorithm has $O(d^3z)$ space and $\poly(d,z)$ update time.
%which implies the same bounds for \Cref{alg:case-1}.
Since we need to run $z + 1$ instances of \Cref{alg:case-1} in parallel, 
the space and update time are multiplied by a factor of~$z$.
\end{proof}

@@

@@

\section{Answering Queries}
\label{sec:query}

%Using \Cref{alg:case-1},
In the following, we show how the information maintained 
by \Cref{alg:case-1} can be used to answer queries of this kind:
find two $\alpha$-separated congruent balls of minimum radius
to cover all but at most $z$ points of the stream processed so far.

Our query algorithm is presented in \Cref{alg:query}.
The idea behind the algorithm is as follows.
By our initial assumption about $p_1$ and by invariants (c), 
we know that $B_1$ completely contains $B_1^*$.
%because $B_1 \cap B_2^* = \emptyset$ . 
But for $B_2$, it might be the case that our assumption about $c_2$ 
was incorrect, and therefore, $B_2$ may not completely contain $B_2^*$. 
To overcome this issue, we try all possible candidates for $c_2$,
and compute, for each possible combinations of 
$B_1$, $B_2$, and Buffer, the best solution for 2-center with $z$ outliers
using the \textproc{MinCover} function presented in \Cref{alg:compute-min}.

\begin{algorithm}
\caption{\sc Query} 
\algsetup{indent=1.5em}
\label{alg:query}
\begin{algorithmic}[1]
	%\STATE solutions $\gets \set{\Call{MinCover}{B_1, B_2, \Buffer}}$
	\STATE solutions $\gets \set{}$

	\STATE candidates $\gets \Buffer \cup B_2$
	\IF {$\card{B_2} \ge \dz$}
		\STATE candidates $\gets \set{c_2} \cup \Buffer $
		\STATE $B_1 \gets B_1 \cup B_2$
	\ENDIF

	\STATE $\delta_0 = \delta$
	\FOR{$c \in$ candidates}
		\STATE $\rc  \gets \frac{2}{\alpha} \len{c_1 c}$
		\STATE $B'_1 \gets B_1$, $\delta \gets \max \{ \delta_0, r \}$
		\STATE $B'_2 \gets \emptyset$, $\Buffer' \gets \emptyset$ 
		\FOR{$p \in$ candidates}
			\IF {\NOT (\Call{AddToB$'_1$}{$p$} \OR \Call{AddToB$'_2$}{$p$})}
				\STATE add $p$ to $\Buffer'$
			\ENDIF
		\ENDFOR
		\STATE add \Call{MinCover}{$B'_1$, $B'_2$, $\Buffer'$} to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{\sc MinCover($B_1, B_2, \text{\rm Buffer}$)} 
\algsetup{indent=1.5em}
\label{alg:compute-min}
\begin{algorithmic}[1]
%\Function{MinCover}{$B_1$, $B_2$, Buffer}
	\STATE solutions $\gets \set{}$
	\FOR{$k \gets 0, \dots,(z - \card{\mbox{Buffer}})$}
		\STATE $r_1 \gets \Call{1-Center}{B_1, k}$
		\STATE $r_2 \gets \Call{1-Center}{B_2, z - \card{\mbox{Buffer}} - k}$
		\STATE add $\max \set{r_1, r_2}$ to solutions
	\ENDFOR
	\RETURN $\min\set{\text{solutions}}$
%\EndFunction
\end{algorithmic}
\end{algorithm}

%Now, we show that the query algorithm correctly considers all possible candidates for $c_2$.
Let $C$ denote the set of candidates for $c_2$.
By invariant~(c), we know that $B_1 \cap B_2^* = \emptyset$. 
Therefore, there exists a point in $(B_2 \cup \Buffer) \cap B_2^*$,
and hence, $C$ is $(B_2 \cup \Buffer) \cap B_2^*$ in general.
%We now use the following lemma.
%In this general case, our candidates $C$ for $c_2$ is set $(B_2 \cup \Buffer) \cap B_2^*$.
However, when $|B_2| \geq \dz$, 
we show in the following lemma that $(\{c_2\} \cup \Buffer) \cap B_2^* \neq \emptyset$. 
Therefore, if $|B_2| \ge \dz$, 
we only need to consider $\{c_2\} \cup \Buffer$ as candidates for $C$.
%our candidates for $C$ are only $\{c_2\} \cup \Buffer$.

@@

\noindent
Our query algorithm works as follows. 
For each candidate point $c \in C$, \Cref{alg:query} 
constructs $B'_1(c_1, \max \{\delta, \radius{c} \})$ and $B'_2(c, \radius{c})$. 
If the candidate $c$ equals the current $c_2$, then, $B_1 = B'_1$. 
Since $\radius{c_2} \leq r \leq \delta/2$ by invariant (b1) and $B'_2 \subset B_2$,
we do not need to construct any new set. 
For $c \neq c_2$, we know that $B_1 \subset B'_1$, and hence,
we just need to see which points in $\Buffer \cap B_2$ are located in $B'_1$.
When $|B_2| \geq \dz$ and $c \neq c_2$, then it means that assumed $c_p \not \in B_2^*$. 
Therefore, by Lemma \ref{lem:B2dz}, $B_2$ can be added to $B'_1$ without contradicting invariant (c). 
So in this case, we just need to see which points of $\Buffer$ must be added to $B'_1$. 
\Cref{alg:query} use functions \textproc{AddToB$'_1$} and \textproc{AddToB$'_2$} for adding a point to $B'_1$ and $B'_2$ respectively. 
The definition of these functions are the same as \textproc{AddToB$_1$} and \textproc{AddToB$_2$},
except that, they add point to $B'_i$ instead of $B_i$ ($i = 1, 2$).

Since \Cref{alg:query} considers all valid candidates for $c$, 
at least for one $c^* \in C$, we have $c^* \in B_2^*$. 
We call the corresponding $B'_1$ and $B'_2$ by $B''_1$ and $B''_2$. 
Since by Observation \ref{obs:c+4},
$1 \leq \frac{\len{c_1 c^*}}{\alpha^*} < \frac{\len {c_1 c^*}}{\alpha r^*}$, 
we have $2r^* \leq \radius{c^*}$, and therefore,
$B^*_1 \subset B''_1$ and $B^*_2 \subset B''_2$ by Lemma~\ref{lem:2r}.
On the other hand, since the distance of the new points added to $B''_1$ compare to $B_1$ is less than $\len{c_1 p} \leq \max \{\delta, \radius{c^*} \}$, 
we have by invariant~(c) that  $B''_1 \cap B_2^* = \emptyset$.
As a result, $B''_1$ (resp., $B''_2$) completely covers $B_1^*$  (resp., $B_2^*$), 
and points in $\Buffer$ are certainly outlier.
The only unknown part is that \Cref{alg:query} does not know how many outliers are in $B''_1$ and $B''_2$.
Therefore, \Cref{alg:compute-min} tries all possible cases and choose the one with the minimum radius.

@@

\REM{ }% to be added in the camera-ready version
\paragraph{Acknowledgement}
The authors would like to thank Kiana Ehsani and Sahand Mozaffari 
for their thoughtful discussions, 
and for their help to improve the quality of this paper.
